### 2022-06-14 paper/paper.tex
paper: final manuscript

\pagenumbering{roman}
\addcontentsline{toc}{chapter}{Abstract}
\begin{center}
	\LARGE\textbf{Abstract}
\end{center}

Typing tests list out words for the user to repeat. These tests measure metrics
that quantify the user's ability to type. However, conventional typing metrics
do not measure correct finger placement. This aspect of typing may affect the
user's health. It is also a crucial aspect of keyboard typing education. As
such, a method to identify which finger is used to press which key is
beneficial. This paper introduces a technique to achieve this by developing a
finger-key identification module that utilizes computer vision algorithms to
detect the keys in a keyboard, and a ready-made machine learning solution to
track fingers while typing. This module was successful in finger-key
identification with an accuracy of 99.58\% in a dataset of 938 keypresses. The
module also had an average finger-key identification time of 0.083 seconds,
which allows the module to perform real-time finger-key identification for
typing speeds up to 143.030 Words per Minute which is well beyond the median
typing speed of the population. Furthermore, two metrics for correct finger
placement were defined and were implemented in a proof-of-concept trainer: (1)
Finger Placement Accuracy which computes the number of keys pressed using the
correct finger in a typing test sequence, and (2) Historical Finger Placement
Accuracy which evaluates the user's ability to press a certain key with the
correct finger over multiple typing test sequences.

@@

\begin{figure}[H]
	\centering
	\includegraphics[width=0.7\textwidth]{60.png}
	\caption{A 60\% keyboard in \ac{ansi} layout.}
	\label{fig:60}
	\centering
\end{figure}

\section{Significance of the Research}
This research is beneficial for all users of physical keyboards. These include a
majority of the population as there are a lot of professions that heavily rely
on keyboards. Examples include developers, physicians, educators, and
accountants. By having better ergonomics while typing --- by touch typing
correctly --- wrist injuries can be prevented, and typing speed may be
increased.

@@

Examples of projects that utilized MediaPipe and its solutions include a
real-time hand gesture recognition system by \textcite{mediapipe-sample-1} which
utilized MediaPipe Hands. Another example is by \textcite{mediapipe-sample-2}
which utilized MediaPipe Hands to track movements during pen spinning
competitions. \textcite{mediapipe-sample-3} developed a real-time sign language
recognition program using MediaPipe Hands.

@@

A toolbox offered by MATLAB is the Computer Vision Toolbox which contains
algorithms, and functions for use in the development of computer vision, 3D
vision, and video processing systems. By using the available algorithms in the
toolbox, such as YOLOv2, and ACF, hand detection and gesture recognition is made
possible in the platform \parencite{matlab}.

@@

This section provides an overview of the development and testing of a real-time
finger-key identification module. First, the experimental setup used for data
collection will be presented. Following that, the algorithm used to implement
real-time finger-key identification will be discussed. The process of training
and testing of the developed algorithm comes next. Afterwards, the development
and design of the proof of concept trainer will be described. Finally, this
section introduces two metrics that were developed to quantify correct finger
placement during typing tests.

@@

The specific camera used was a Logitech C920. It is a 3 mega pixel webcam that
is capable of capturing color video in 1080p/30fps and 720p/30fps with a
diagonal field of view of 78\degree. For the setup, the camera captured the
video in 480p/30fps. This camera has a universal mounting clip that allows the
camera to be correctly positioned within the experimental setup
\parencite{logitech}.

@@

\subsubsection{Real-Time Finger-Key Identification}
With each keypress only requiring 0.083s for finger-key identification, it can
be gathered that the maximum \ac{cps} which the module can still perform
real-time finger-key identification is 11.9200 \ac{cps}, based on
Equation~\ref{eq:s}. It follows that the maximum \ac{wpm} that still allows for
real-time finger-key identification is 143.040 \ac{wpm}, based on
Equation~\ref{eq:cps}. This is well above the median of 37.5 \ac{wpm} on Keybr.
The module can also perform finger-key identification for 90.99\% of the
community members of Monkeytype with the speed role.

@@

The module is fast enough for real-time finger-key identification up to typing
speeds of 143.040 \ac{wpm}. This is above the median typing speed of 37.5
\ac{wpm} on Keybr \parencite{keybr}. This also handles 90.99\% of the 12678
members with the speed role on MonkeyType \parencite{monkey-stats}.

>>>>

\pagenumbering{roman}

\addcontentsline{toc}{chapter}{Abstract}
\begin{center}
	\LARGE\textbf{Acknowledgment}
\end{center}

I would like to express my deepest appreciation for the guidance that my Special
Project Adviser, Dhong Fhel K. Gom-os, has given me throughout the duration of
the project. His technical knowledge has been a key factor to the success of the
finger-key identification module. The paper would not have been possible without
his unwavering support and commitment.

I also greatly appreciate the advice that Prof. Aileen Joan O. Vicente and Ms.
Sram J Isle L. Santillan have given me. The constructive criticism that I
received was instrumental in giving me a clear direction on where I want to take
this research.

I am particularly grateful to Jack Bartnik, the developer of Monkeytype. The
application that they developed was the main source of inspiration for this
project. I would also like to thank Jack for their generosity in sharing
statistics concerning the typing speed of the users of Monkeytype.

I would like to acknowledge that this research was partially supported by the
Department of Science and Technology --- Science Education Institute.

I give my heartfelt thanks to Ana Janelle H. Ca\~{n}ete for her love and comfort
in these trying times.

To friends, family, and other significant people in my life, thank you for
supporting me thus far.

\newpage
\addcontentsline{toc}{chapter}{Abstract}
\begin{center}
	\LARGE\textbf{Abstract}
\end{center}

Typing tests list out words for the user to repeat. These tests measure metrics
that quantify the user's ability to type. However, conventional typing metrics
do not measure correct finger placement. This aspect of typing may affect the
user's health. It is also a crucial aspect of keyboard typing education. As
such, a method to identify which finger is used to press which key is
beneficial. This paper introduces a technique to achieve this by developing a
finger-key identification module that utilizes computer vision algorithms to
detect the keys in a keyboard, and a ready-made machine learning solution to
track fingers while typing. This module was successful in finger-key
identification with an accuracy of 99.58\% in a dataset of 942 keypresses. The
module also had an average finger-key identification time of 0.083 seconds,
which allows the module to perform real-time finger-key identification for
typing speeds up to 143.030 Words per Minute which is well beyond the median
typing speed of the population. Furthermore, two metrics for correct finger
placement were defined and were implemented in a proof-of-concept trainer: (1)
Finger Placement Accuracy which computes the number of keys pressed using the
correct finger in a typing test sequence, and (2) Historical Finger Placement
Accuracy which evaluates the user's ability to press a certain key with the
correct finger over multiple typing test sequences.

@@

\begin{figure}[H]
	\centering
	\includegraphics[width=0.7\textwidth]{60.png}
	\caption{A 60\% keyboard in modified \ac{ansi} layout.}
	\label{fig:60}
	\centering
\end{figure}

\section{Significance of the Research}
This research is beneficial for all users of physical keyboards. These include a
majority of the population as there are a lot of professions that rely heavily
on keyboards. Examples include developers, physicians, educators, and
accountants. By having better ergonomics while typing --- by touch typing
correctly --- wrist injuries can be prevented, and typing speed may be
increased.

@@

Examples of projects that utilized MediaPipe and its solutions include a
real-time hand gesture recognition system by \textcite{mediapipe-sample-1} which
utilized MediaPipe Hands. Another example is by \textcite{mediapipe-sample-2}
who utilized MediaPipe Hands to track movements during pen spinning
competitions. \textcite{mediapipe-sample-3} developed a real-time sign language
recognition program using MediaPipe Hands.

@@

A toolbox offered by MATLAB is the Computer Vision Toolbox which contains
algorithms, and functions for use in the development of computer vision, 3D
vision, and video processing systems. By using the available algorithms in the
toolbox, such as YOLOv2, and ACF, hand detection and gesture recognition are
made possible in the platform \parencite{matlab}.

@@

This section provides an overview of the development and testing of a real-time
finger-key identification module. First, the experimental setup used for data
collection will be presented. Following that, the algorithm used to implement
real-time finger-key identification will be discussed. The process of training
and testing the developed algorithm comes next. Afterward, the development and
design of the proof of concept trainer will be described. Finally, this section
introduces two metrics that were developed to quantify correct finger placement
during typing tests.

@@

The specific camera used was a Logitech C920. It is a 3 megapixel webcam that
is capable of capturing color video in 1080p/30fps and 720p/30fps with a
diagonal field of view of 78\degree. For the setup, the camera captured the
video in 480p/30fps. This camera has a universal mounting clip that allows the
camera to be correctly positioned within the experimental setup
\parencite{logitech}.

@@

\subsubsection{Real-Time Finger-Key Identification}
With each keypress only requiring 0.083s for finger-key identification, it can
be gathered that the maximum \ac{cps} which the module can still perform
real-time finger-key identification is 11.9200 \ac{cps}, based on
Equation~\ref{eq:s}. It follows that the maximum \ac{wpm} that still allows for
real-time finger-key identification is 143.040 \ac{wpm}, based on
Equation~\ref{eq:cps}. This is well above the median of 37.5 \ac{wpm} on Keybr.
The module can also perform real-time finger-key identification for 90.99\% of
the community members of Monkeytype with the speed role.

@@

The module is fast enough for real-time finger-key identification up to typing
speeds of 143.040 \ac{wpm}. This is above the median typing speed of 37.5
\ac{wpm} on Keybr \parencite{keybr}. This also handles 90.99\% of the 12678
members with the speed role on Monkeytype \parencite{monkey-stats}.
